---
title: "Matching parcels to farms investigation"
author: "Selen Ozdogan"
date: "4/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sf)
library(lwgeom)
library(ggplot2)
library(rgeos)
library(raster)
library(rgdal)
library(polylabelr)
library(janitor)
library(terra)
library(dplyr)
library(magrittr)
library(ggpubr)
library(plotly)

gc()
setwd("~/Dropbox/land_ineq_degradation")
sfiles <- "georeferenced_data/cadastre/FR"
cfiles <- "georeferenced_data/gadm/gadm36_FRA_shp"
t = 2015

# Read matched file created by R
matched_raw <- clean_names(read_sf(file.path(sfiles, t, "Rtests/matched.shp")))

# Drop 99% of the data to work faster
matched_raw$drop <- runif(nrow(matched_raw), 0, 100)
matched_raw <- matched_raw[matched_raw$drop > 99,]

# Clean missing and duplicated values
matched_na <- as.numeric(sum(is.na(matched_raw$farm_area)))
matched_dup <- as.numeric(sum(duplicated(matched_raw$id_parcel)))

matched <- matched_raw[!is.na(matched_raw$farm_area),]
matched <- matched[!duplicated(matched$id_parcel),]

# Calculate some stats
matched_count_farm <- as.numeric(length(unique(matched$id_ilot)))
matched_count_parc <- as.numeric(length(unique(matched$id_parcel)))
matched_total_farmarea <- sum(matched$parc_area)

# Collapse to farm level
matched %<>%
  dplyr::select(id_ilot, surf_parc, parc_area, farm_area) %>%
  mutate(id_ilot = as.numeric(id_ilot), surf_parc = as.numeric(surf_parc),
         parc_area = as.numeric(parc_area), farm_area = as.numeric(farm_area)) %>%
  arrange(id_ilot) %>%
  group_by(id_ilot) %>%
  summarise(sum_parc_area = sum(parc_area), sum_surf_parc = sum(surf_parc),
            farm_area = mean(farm_area), count_parc = n())

# Calculate percent difference between parcel area and surface sum and farm area
matched$diff_area = ((matched$sum_parc_area - matched$farm_area)/matched$farm_area)*100
matched$diff_surf = ((matched$sum_surf_parc - (matched$farm_area/10000))/(matched$farm_area/10000))*100

# Collapse to canton level
gadm4 <- clean_names(read_sf(file.path(cfiles, "gadm36_FRA_4.shp")))
gadm4 <- st_transform(gadm4, crs(matched))
matched_canton <- st_join(gadm4, matched, left = TRUE)

# Clean missing values
matched_canton_na_surf <- as.numeric(sum(is.na(matched_canton$diff_surf)))
matched_canton_na_area <- as.numeric(sum(is.na(matched_canton$diff_area)))

matched_canton <- matched_canton[!is.na(matched_canton$diff_surf),]
matched_canton <- matched_canton[!is.na(matched_canton$diff_area),]

matched_canton_count_farm <- as.numeric(length(unique(matched_canton$id_ilot)))
matched_canton_count_parc <- as.numeric(sum(matched_canton$count_parc))

# Calculate more stats
matched_sum_parc_surf <- sum(matched_canton$sum_surf_parc)
matched_avg_parc_surf <- mean(matched_canton$sum_surf_parc)

matched_sum_parc_area <- sum(matched_canton$sum_parc_area)
matched_avg_parc_area <- mean(matched_canton$sum_parc_area)

# Take average diffs by canton
matched_cavg <- matched_canton %>%
  group_by(gid_4) %>%
  summarise(avg_diff_area = mean(diff_area), avg_diff_surf = mean(diff_surf),
            count_parc = sum(count_parc), count_farm = n())

# Two more stats
matched_mean_diffsurf <- mean(matched_cavg$avg_diff_surf)
matched_mean_diffarea <- mean(matched_cavg$avg_diff_area)

# Create only negative and only positive subsets for plots
matched_cavg_apos <- matched_cavg[matched_cavg$avg_diff_area > 0,]
matched_cavg_aneg <- matched_cavg[matched_cavg$avg_diff_area < 0,]
matched_cavg_spos <- matched_cavg[matched_cavg$avg_diff_surf > 0,]
matched_cavg_sneg <- matched_cavg[matched_cavg$avg_diff_surf < 0,]

####

# Read matched file created by PyQGIS
matchedo_raw <- clean_names(read_sf(file.path(sfiles, t, "parcelles_matched_to_ilots.shp")))

# Drop 99% of the data to work faster
matchedo_raw$drop <- runif(nrow(matchedo_raw), 0, 100)
matchedo_raw <- matchedo_raw[matchedo_raw$drop > 99,]

# Clean missing and duplicated values
matchedo_na <- as.numeric(sum(is.na(matchedo_raw$area)))
matchedo_dup <- as.numeric(sum(duplicated(matchedo_raw$id_parcel)))

matchedo <- matchedo_raw[!is.na(matchedo_raw$area),]
matchedo <- matchedo[!duplicated(matchedo$id_parcel),]

# Calculate some stats
matchedo_count_farm <- as.numeric(length(unique(matchedo$id_ilot)))
matchedo_count_parc <- as.numeric(length(unique(matchedo$id_parcel)))
matchedo_total_farmarea <- sum(matchedo$area)

# Collapse to farm level
matchedo %<>%
  dplyr::select(id_ilot, surf_parc, area) %>%
  mutate(id_ilot = as.numeric(id_ilot), surf_parc = as.numeric(surf_parc),
         area = as.numeric(area)) %>%
  arrange(id_ilot) %>%
  group_by(id_ilot) %>%
  summarise(sum_surf_parc = sum(surf_parc), farm_area = mean(area), count_parcel = n())

# Calculate percent difference between parcel surface sum and farm area
matchedo$diff_surf = ((matchedo$sum_surf_parc - (matchedo$farm_area/10000))/(matchedo$farm_area/10000))*100

# Collapse to canton level
gadm4 <- st_transform(gadm4, crs(matchedo))
matchedo_canton <- st_join(gadm4, matchedo, left = TRUE)

# Clean missing values
matchedo_canton_na <- as.numeric(sum(is.na(matchedo_canton$diff_surf)))
matchedo_canton <- matchedo_canton[!is.na(matchedo_canton$diff_surf),]

# Calculate more stats
matchedo_sum_parc_surf <- sum(matchedo_canton$sum_surf_parc)
matchedo_avg_parc_surf <- mean(matchedo_canton$sum_surf_parc)

matchedo_canton_count_farm <- as.numeric(length(unique(matchedo_canton$id_ilot)))
matchedo_canton_count_parc <- as.numeric(sum(matchedo_canton$count_parcel))

# Take average diffs by canton
matchedo_cavg <- matchedo_canton %>% 
  group_by(gid_4) %>% 
  summarise(avg_diff_surf = mean(diff_surf), count_farm = n()) 

# One more stats
matchedo_mean_diffsurf <- mean(matchedo_cavg$avg_diff_surf)

# Create only negative and only positive subsets for plots
matchedo_cavg_spos <- matchedo_cavg[matchedo_cavg$avg_diff_surf > 0,]
matchedo_cavg_sneg <- matchedo_cavg[matchedo_cavg$avg_diff_surf < 0,]
```

# Matched file created by R

### Some stats:

* In the R and QGIS matched files, respectively:
    + farm_area variable has `r matched_na` and `r matchedo_na` missing values.
    + id_parcel variable has `r matched_dup` and `r matchedo_dup` duplicated values.
    + Total farm area is `r matched_total_farmarea` and `r matchedo_total_farmarea`.
    + There are `r matched_count_farm` and `r matchedo_count_farm` farms.
    + There are `r matched_count_parc` and `r matchedo_count_parc` parcels.


* In the R and QGIS matched files aggregated by canton:
    + diff_surf variable has `r matched_canton_na_surf` and `r matchedo_canton_na` missing values.
    + Average difference in surface is `r matched_mean_diffsurf` and `r matchedo_mean_diffsurf`.
    + Average difference in area is `r matched_mean_diffarea` in the R file.
    + Total parcel area is `r matched_sum_parc_area` and average parcel area is `r matched_avg_parc_area` in the R file.
    + Total parcel surface is `r matched_sum_parc_surf` and `r matchedo_sum_parc_surf`.
    + Average parcel surface is `r matched_avg_parc_surf` and `r matchedo_avg_parc_surf`.
    + There are `r matched_canton_count_farm` and `r matchedo_canton_count_farm` farms.
    + There are `r matched_canton_count_parc` and `r matchedo_canton_count_parc` parcels.

## Farm area vs. parcel area percent difference

### All values
```{r}
aall <- ggplot(matched_cavg) + geom_sf(aes(fill = avg_diff_area), lwd = 0)
ggplotly(aall)
```

### Only positive values
```{r}
apos <- ggplot(matched_cavg_apos) + geom_sf(aes(fill = avg_diff_area), lwd = 0)
ggplotly(apos)
```

### Only negative values
```{r}
aneg <- ggplot(matched_cavg_aneg) + geom_sf(aes(fill = avg_diff_area), lwd = 0)
ggplotly(aneg)
```

# Farm area vs. parcel surface percent difference

### All values
```{r}
sall <- ggplot(matched_cavg) + geom_sf(aes(fill = avg_diff_surf), lwd = 0)
ggplotly(sall)
```

### Only positive values
```{r}
spos <- ggplot(matched_cavg_spos) + geom_sf(aes(fill = avg_diff_surf), lwd = 0)
ggplotly(spos)
```

### Only negative values
```{r}
sneg <- ggplot(matched_cavg_sneg) + geom_sf(aes(fill = avg_diff_surf), lwd = 0)
ggplotly(sneg)
```

# Matched file created by QGIS

## Farm area vs. parcel surface percent difference

### All values
```{r}
orig_sall <- ggplot(matchedo_cavg) + geom_sf(aes(fill = avg_diff_surf), lwd = 0)
ggplotly(orig_sall)
```

### Only positive values
```{r}
orig_spos <- ggplot(matchedo_cavg_spos) + geom_sf(aes(fill = avg_diff_surf), lwd = 0)
ggplotly(orig_spos)
```

### Only negative values
```{r}
orig_sneg <- ggplot(matchedo_cavg_sneg) + geom_sf(aes(fill = avg_diff_surf), lwd = 0)
ggplotly(orig_sneg)
```

## Exluding extreme positive values (> 100%)

```{r, include=FALSE}
matchedo_cavg_100 <- matchedo_cavg[matchedo_cavg$avg_diff_surf <= 100,]
matchedo_cavg_spos_100 <- matchedo_cavg_spos[matchedo_cavg_spos$avg_diff_surf <= 100,]
```

### All values
```{r}
orig_sall_100 <- ggplot(matchedo_cavg_100) + geom_sf(aes(fill = avg_diff_surf), lwd = 0)
ggplotly(orig_sall_100)
```

### Only positive values
```{r}
orig_spos_100 <- ggplot(matchedo_cavg_spos_100) + geom_sf(aes(fill = avg_diff_surf), lwd = 0)
ggplotly(orig_spos_100)
```
